For submission to the Journal of Integrated Pest Management

### 2021-03-09 ofv_fife_2021.Rmd
Revised according to Dr. Martini's suggestions

### Body Text
*Orchid fleck dichorhavirus* (OFV) is the type species for the genus *Dichorhavirus*, a group of plant-infecting, bacilliform, nuclear rhabdoviruses composed of two segments of single-stranded, negative-sense RNA [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Other members of this genus are: *Citrus chlorotic spot dichorhavirus*, *Citrus leprosis N dichorhavirus*, *Clerodendrum chlorotic spot dichorhavirus* and *Coffee ringspot dichorhavirus* [@Dietzgen2018a]. Flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae) are the only known vector for dichorhaviruses [@Maeda1998]. These mites transmit OFV in a persistent propagative manner [@Kondo2003].

OFV-infected plants exhibit various symptoms dependent on the variety of plant infected as well as the strain of the virus [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns (*`r citef("fig_1")`*).
<!-- These lesions are thought to be associated with the feeding sites of the mite vector [@cite]. Also, is it probably a sign of RNAi plant defenses: see Hull Plant Virology -->

@@

On June 2020, chlorotic ringspot symptoms were observed on the liriopogon Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in the landscape of Leon County, Florida (*`r citef("fig_1")`*). Liropogons belong to the plant family Asparagaceae, subfamily Nolinoideae, which includes various monocotyledonous lilliod plants [@Chase2009]. Liriopogons are native to southeastern Asia, encompassing the closely-related genera *Liriope* and *Ophiopogon*, which are characterized by evergreen, long grass-like leaves, and erect flowers which produce black, purple or blue, berry-like fruits, depending on the cultivar [@Lattier2014; @Fantz2015]. These characteristics have given rise to various common names, including: monkey grass, Aztec grass, lilyturf, mondo grass, etc. (*`r citet("tab_1")`*). The genus *Liriope* contains eight species, four of which are commonly cultivated, while the genus *Ophiopogon* consists of about 65 species, with only a handful are available for horticulture [@Lattier2014; @Nesom2010]. Liripogons are considered the most important ground cover sold by the nursery industry in southeastern US [@Mcharo2003].

Plant leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. The plant samples were tested for a wide range of pathogens, including **[Waiting on Dr. Iriarte's response]** but the results were inconclusive. The infected materials were subsequently sent to the Florida Department of Agriculture and Consumer Services (FDACS). The presence of *Orchid fleck dichorhavirus* was confirmed via generic one step conventional RT-PCR with R2-Dicho-GF and R2-Dicho-GR primers, amplifying ~800 nt of L-gene (RNA2) amplicon from a leaf sample of infected *Liriope* [@Kubo2009b]. RT-qPCR confirmed OFV and the RT-PCR product was sequenced using Sanger sequencing and High Throughput Sequencing [@Kondo2006]. Sanger sequencing of RT-PCR amplicons shared 98 \% nucleotide identity with OFV strains which infect orchids (GenBank Accession numbers: AB244418 and MK522807) [@Kondo2006; @Kondo2014; @Cook2019]. Further samples were taken from various symptomatic cultivars of *Liriope* spp., *Ophiopogon* spp., as well as *Aspidistra elatior* Blume (Asparagales: Nolinoideae) during subsequent visits to the initial site of collection as well as other locations in Leon county. Samples were tested via RT-PCR at the NFREC (*`r citet("tab_1")`*) to confirm the presence of OFV.

Further surveys of Nolinoideae in Florida have revealed more sites with symptomatic plants in both Leon and Alachua counties. In addition to the discovery of the OFV-infected plants, we collected mites from symptomatic plants in Leon county which were observed with phase contrast microscopy. We encountered both eriophyoid mites and flat mites which were identified as *Brevipalpus californicus* (Banks) sensu lato, an identity which was confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. *Brevipalpus* mites have been previously associated with OFV [@BeltranBeltran2020; @GarciaEscamilla2018; @Dietzgen2018b] and similar diseases [@Kitajima2010] and are known to feed on a large variety of economically-important plants [@Childers2003a; @Akyazi2017]. Unfortunately, the *Brevipalpus* mite species complex is known to contain cryptic species [@Childers2011] whose identification can be improved with molecular methods as well as more advanced microscopy techniques, such as low-temperature scanning electron microscopy (LT-SEM) [@Skoracka2015; @Leon2010]. With that in mind, we sent additional samples of the mites to the USDA-ARS in Beltsville to observe the mites with LT-SEM techniques, which agreed with both prior identifications of *B. californicus* s.l. (*`r citef("fig_2")`*).

@@

Detecting OFV in Florida represents a concern for horticulturalists who grow susceptible plants of economic importance, including orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagales spp. which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including the famous Ghost Orchid, *Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe. Furthermore, OFV represents an obstacle to overcome for the burgeoning interest in cultivating *Vanilla* in southern Florida [@Chambers2019]. Lastly, some OFV isolates are known to be involved with nuclear types of citrus leprosis [@Roy2015a], which may be a cause for concern for the citrus industry. In fact, @Kitajima2011a found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis, which are closely related to OFV strains [@Roy2013a]. *B. californicus* and *B. yothersi* are both known vectors of Dichorhaviruses (OFV) and Cileviruses (Citrus Leprosis) [@BeltranBeltran2020; @Kondo2003; @Knorr1968] and *B. obovatus* is a suspected vector as well [@Childers2003a]. All three mite species/complexes are present in Florida [@Childers2003a; @Akyazi2017] (*`r citef("fig_3")`*); therefore, it is critical to monitor the spread of Orchid Fleck Virus and its mite vector(s) to determine what risk this virus represents for plants in Florida and the surrounding regions.

### Brief Section on Control Options

The predatory mite *Galendromus helveolus* (Acari: Phytoseiidae) have been reported to feed on eggs, larvae and nymphs of *B. californicus* [@Chen2006].

The entomopathogenic fungus *Metarhizium anisopliae* var. *acridum* has been shown to be pathogenic to *B. phoenicis* [@Magalhaes2005], we have encountered similar fungal infections

@@

### Figures and Tables

```{r, fig_1, warning=FALSE}
grid::grid.raster(tiff::readTIFF( "images/fig_1.tif"))
```
`r fig_nums(name = "fig_1", caption = "Variety of symptoms expressed by plants infected with Orchid fleck dichorhavirus: (a) ringspot symptoms on *Liriope gigantea* (b) chlorotic flecking on *Aspidistra elatior*.")`

```{r, fig_2, warning=FALSE}
grid::grid.raster(tiff::readTIFF( "images/b_cali.tif"))
```
`r fig_nums(name = "fig_2", caption = "LT-SEM images of *Brevipalpus californicus* sensu lato displaying various characters used for identification [@Baker1987; @Beard2012] (a) Dorsum  (b) Lateral view (c) Venter (d) Close up of distal end of leg 2, with arrows indicating paired solenidia, characteristic of the genus *Brevipalpus* (e) Enlargement of the microplates of the mite cerotegument (f) Dorsal view of the distal portion of mite abdomen (g) Dorsal view of the mite rostrum (h) Ventral view of mite rostrum, observe 3 distal setae. LT-SEM images provided by Dr. Gary Bauchan, USDA-ARS 2021")`

```{r, fig_3, warning=FALSE}
grid::grid.raster(tiff::readTIFF( "images/other_mites.tif"))
```
`r fig_nums(name = "fig_3", caption = "Florida is home to other common pest species of *Brevipalpus* which are potential vectors of *Orchid fleck dichorhavirus*: (a) *B. yothersi*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal. SEM images provided by Dr. Ron Ochoa, USDA-ARS 2021")`

@@

>>>>

\pagebreak
*Orchid fleck dichorhavirus* (OFV) is the type species for the genus *Dichorhavirus*, a group of plant-infecting, bacilliform, nuclear rhabdoviruses composed of two segments of single-stranded, negative-sense RNA [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Other members of this genus are: *Citrus chlorotic spot dichorhavirus*, *Citrus leprosis N dichorhavirus*, *Clerodendrum chlorotic spot dichorhavirus* and *Coffee ringspot dichorhavirus* [@Dietzgen2018a]. Flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae) are the only known vector for dichorhaviruses [@Maeda1998]. These mites transmit OFV in a persistent propagative manner [@Kondo2003].

OFV-infected plants exhibit various symptoms dependent on the variety of plant infected as well as the strain of the virus [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns (`r citef("fig_1")`).
<!-- These lesions are thought to be associated with the feeding sites of the mite vector [@cite]. Also, is it probably a sign of RNAi plant defenses: see Hull Plant Virology -->

@@

During June 2020, chlorotic ringspot symptoms were observed on the liriopogon Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in the landscape of Leon County, Florida (`r citef("fig_1")`). Liropogons belong to the plant family Asparagaceae, subfamily Nolinoideae, which includes various monocotyledonous lilliod plants [@Chase2009]. Liriopogons are native to southeastern Asia, encompassing the closely-related genera *Liriope* and *Ophiopogon*, which are characterized by evergreen, long grass-like leaves, and erect flowers which produce black, purple or blue, berry-like fruits, depending on the cultivar [@Lattier2014; @Fantz2015]. These characteristics have given rise to various common names, including: monkey grass, Aztec grass, lilyturf, mondo grass, etc. (*`r citet("tab_1")`*). The genus *Liriope* contains eight species, four of which are commonly cultivated, while the genus *Ophiopogon* consists of about 65 species, with only a handful are available for horticulture [@Lattier2014; @Nesom2010]. Liripogons are considered the most important ground cover sold by the nursery industry in southeastern US [@Mcharo2003].

Plant leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. The plant samples were tested for a wide range of common plant pathogens, but the results were inconclusive. The infected materials were subsequently sent to the Florida Department of Agriculture and Consumer Services (FDACS). The presence of *Orchid fleck dichorhavirus* was confirmed via generic one step conventional RT-PCR with R2-Dicho-GF and R2-Dicho-GR primers, amplifying ~800 nt of L-gene (RNA2) amplicon from a leaf sample of infected *Liriope* [@Kubo2009b]. RT-qPCR confirmed OFV and the RT-PCR product was sequenced using Sanger sequencing and High Throughput Sequencing [@Kondo2006]. Sanger sequencing of RT-PCR amplicons shared 98 \% nucleotide identity with OFV strains which infect orchids (GenBank Accession numbers: AB244418 and MK522807) [@Kondo2006; @Kondo2014; @Cook2019]. Further samples were taken from various symptomatic cultivars of *Liriope* spp., *Ophiopogon* spp., as well as *Aspidistra elatior* Blume (Asparagales: Nolinoideae) during subsequent visits to the initial site of collection as well as other locations in Leon county. Samples were tested via RT-PCR at the NFREC (*`r citet("tab_1")`*) to confirm the presence of OFV.

Further surveys of Nolinoideae in Florida have revealed more sites with symptomatic plants in both Leon and Alachua counties. In addition to the discovery of the OFV-infected plants, we collected mites from symptomatic plants in Leon county which were observed with phase contrast microscopy. We encountered both eriophyoid mites and flat mites which were identified as *Brevipalpus californicus* (Banks) sensu lato, an identity which was confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. *Brevipalpus* mites have been previously associated with OFV [@BeltranBeltran2020; @GarciaEscamilla2018; @Dietzgen2018b] and similar diseases [@Kitajima2010] and are known to feed on a large variety of economically-important plants [@Childers2003a; @Akyazi2017]. Unfortunately, the *Brevipalpus* mite species complex is known to contain cryptic species [@Childers2011] whose identification can be improved with molecular methods as well as more advanced microscopy techniques, such as low-temperature scanning electron microscopy (LT-SEM) [@Skoracka2015; @Leon2010]. With that in mind, we sent additional samples of the mites to the USDA-ARS in Beltsville to observe the mites with LT-SEM techniques, which agreed with both prior identifications of *B. californicus* s.l. (`r citef("fig_2")`).

@@

### Control methods for *Brevipalpus* mites

Reducing the populations of the *Brevipalpus* mite vector may be able to reduce the spread of OFV by infective mites, but infected plants are not able to recover from the virus and must be rouged or removed. Rouging plants is a direct and efficient option to remove mite and virus populations quickly but increases in difficultly with the number of plants involved. Once infected plants have been removed, the objective will be to prevent re-infestation of the *Brevipalpus* vector via conventional pest management methods.

Most research on the integrated management of mites has been catered towards controlling various species of spider mites (Acari: Tetranychidae) [@Leeuwen2015], but many of those methods can be applied to *Brevipalpus* spp. as well:

Chemical control via synthetic acaricides is commonly used to control *Brevipalpus* spp. [@Leeuwen2015], but acaricides have some drawbacks. Many products control a broad class of arthropods, but often they do so indiscriminately, killing beneficial insects as well as pests [@Suckling2013]. Control strategies which rely heavily on chemical control have the potential to encourage pesticide resistance in mite populations and pesticide resistance has been reported in various *Brevipalpus* populations [@Alves2000; @Campos2002; @Omoto2000]. In addition, it is important to consider the interactions which may occur between different chemical applications or tank mixes [@Vechia2018]. Lastly, application costs of chemical controls can be large and increase the cost of production [@Rodrigues2000a], nevertheless chemical applications are valuable tools used to consider against *Brevipalpus*.

Various biological control methods have also been studied, but their use has not been widely adopted [@Messing2017]. Even so, some studies have found biological control methods which could be used specifically for *Brevipalpus* mites: The predatory mite *Galendromus helveolus* (Chant) (Acari: Phytoseiidae) has been reported to feed on eggs, larvae and nymphs of *B. californicus* [@Chen2006] and studies of *Amblyseius largoensis* (Muma) (Acari: Phytoseiidae) demonstrated its ability to suppress populations of *B. yothersi* on citrus plants [@Argolo2020]. In addition, the entomopathogenic fungus *Metarhizium anisopliae* var. *acridum* was found to be pathogenic to *B. phoenicis* [@Magalhaes2005]. In our own surveys, we have found mites which succumbed to fungal infections (`r citef("fig_3")`); however the fungal species was not determined.

### Conclusion

Detecting OFV in Florida represents a concern for horticulturalists who grow susceptible plants of economic importance, including orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagales spp. which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including the famous Ghost Orchid, *Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe. Furthermore, OFV represents an obstacle to overcome for the burgeoning interest in cultivating *Vanilla* in southern Florida [@Chambers2019]. Lastly, some OFV isolates are known to be involved with nuclear types of citrus leprosis [@Roy2015a], which may be a cause for concern for the citrus industry. In fact, @Kitajima2011a found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis, which are closely related to OFV strains [@Roy2013a]. *B. californicus* and *B. yothersi* are both known vectors of Dichorhaviruses (OFV) and Cileviruses (Citrus Leprosis) [@BeltranBeltran2020; @Kondo2003; @Knorr1968] and *B. obovatus* is a suspected vector as well [@Childers2003a]. All three mite species/complexes are present in Florida [@Childers2003a; @Akyazi2017] (`r citef("fig_4")`); therefore, it is critical to monitor the spread of Orchid Fleck Virus and its mite vector(s) to determine what risk this virus represents for plants in Florida and the surrounding regions.

@@

### Table

@@

### Figure captions
`r fig_nums(name = "fig_1", caption = "Variety of symptoms expressed by plants infected with Orchid fleck dichorhavirus: (a) ringspot symptoms on *Liriope gigantea* (b) chlorotic flecking on *Aspidistra elatior*.")`

`r fig_nums(name = "fig_2", caption = "LT-SEM images of *Brevipalpus californicus* sensu lato displaying various characters used for identification [@Baker1987; @Beard2012] (a) Dorsum  (b) Lateral view (c) Venter (d) Close up of distal end of leg 2, with arrows indicating paired solenidia, characteristic of the genus *Brevipalpus* (e) Enlargement of the microplates of the mite cerotegument (f) Dorsal view of the distal portion of mite abdomen (g) Dorsal view of the mite rostrum (h) Ventral view of mite rostrum, observe 3 distal setae. LT-SEM images provided by Dr. Gary Bauchan, USDA-ARS 2021")`

`r fig_nums(name = "fig_3", caption = "LT-SEM images of unidentified fungus infecting *Brevipalpus californicus* sensu lato: (a) Infested *B. californicus* adult, dorsal (b) Detail of fungal sporangia. LT-SEM images provided by Dr. Gary Bauchan, USDA-ARS 2021")`

`r fig_nums(name = "fig_4", caption = "Florida is home to other common pest species of *Brevipalpus* which are potential vectors of *Orchid fleck dichorhavirus*: (a) *B. yothersi*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal. SEM images provided by Dr. Ron Ochoa, USDA-ARS 2021")`

### Figures
```{r, fig_1, warning=FALSE}
grid::grid.raster(tiff::readTIFF( "images/fig_1.tif"))
```
```{r, fig_2, warning=FALSE}
grid::grid.raster(tiff::readTIFF( "images/b_cali.tif"))
```
```{r, fig_3, warning=FALSE}
grid::grid.raster(tiff::readTIFF( "images/fungus_mite.tif"))
```
```{r, fig_4, warning=FALSE}
grid::grid.raster(tiff::readTIFF( "images/other_mites.tif"))
```
